---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
<!--
[![Build Status](https://travis-ci.org/hrbrmstr/sergeant.svg)](https://travis-ci.org/hrbrmstr/sergeant) 
![Project Status: Concept - Minimal or no implementation has been done yet.](http://www.repostatus.org/badges/0.1.0/concept.svg)](http://www.repostatus.org/#concept)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/sergeant)](http://cran.r-project.org/web/packages/sergeant) 
![downloads](http://cranlogs.r-pkg.org/badges/grand-total/sergeant)
-->

`sergeant` : Tools to Transform and Query Data with the 'Apache' 'Drill' 'API'

The following functions are implemented:

- `drill_metrics`:	Get the current memory metrics
- `drill_options`:	List the name, default, and data type of the system and session options
- `drill_profiles`:	Get the profiles of running and completed queries
- `drill_query`:	Submit a query and return results
- `drill_stats`:	Get Drillbit information, such as ports numbers
- `drill_status	Get the status of Drill
- `drill_storage`:	Get the list of storage plugin names and configurations
- `drill_threads`:	Get information about threads

### News

- Version 0.1.0 released

### Installation

```{r eval=FALSE}
devtools::install_github("hrbrmstr/sergeant")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r}
library(sergeant)

# current verison
packageVersion("sergeant")

drill_query("SELECT * FROM dfs.`/usr/local/drill/sample-data/nation.parquet`")
```

### Test Results

```{r}
library(sergeant)
library(testthat)

date()

test_dir("tests/")
```

### Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.
